---
title: "Fast food"
output: github_document
---

This was a spontaneous glance at a fast foods data set, taken from [fastfoodnutrition.org](fastfoodnutrition.org).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}

library(readr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(dplyr)
library(wesanderson)
library(ggthemes)

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
theme_set(theme_solarized())


```

# Brands by item count:

Taco Bell appears to offer the highest selection. 

In Lithuania, we only have McDonald's and Subway, with Burger King coming soon. However, I doubt that the Subway franchises over here exceed McDonald's in the items offered.

```{r, fig.width = 10, fig.height= 7}
fastfood %>% 
  group_by(restaurant) %>% 
  summarise(total_items = sum(n())) %>% 
  ggplot(aes(reorder(restaurant, total_items), total_items))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = total_items))+
  labs(x ="Restaurant",
       y ="Total item count")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 8))
```

# Highest calorie item per brand:

The 20 piece Buttermilk Crispy Chicken Tenders could feed someone for an entire day. At what cost, though?

```{r, fig.width = 10, fig.height= 7}
max_cal_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_cal = max(calories))

max_cal_items <- fastfood %>% 
  inner_join(max_cal_numbers,
             by = c("calories" = "max_cal",
                    "restaurant"))

max_cal_items %>% 
  ggplot(aes(restaurant, calories))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")

```

Almost half the calories in this item are from fat.

```{r}

fastfood %>% 
  filter(X1 == 40) %>% 
  mutate(cal_not_fat = calories - cal_fat) %>% 
  select(-c(X1, restaurant, salad)) %>% 
  pivot_longer(cols = -item,
               names_to = "nutrient",
               values_to = "quantity"
               ) %>%
  filter(nutrient %in% c("cal_not_fat", "cal_fat")) %>% 
  mutate(nutrient = factor(nutrient),
         nutrient = fct_recode(nutrient,
           `Calories from fat` = "cal_fat",
           `Calories not from fat` = "cal_not_fat"
         )) %>% 
  ggplot()+
  geom_col(aes(item, quantity, fill = nutrient),
           color = "black")+
  ylab("Calories")+
  scale_fill_manual(values = wes_palette("GrandBudapest1", n = 2))

```

# Lowest calorie item per brand:

Would you really go to a fast food chain for a salad?

Chick-Fil-A's 4 Piece Chicken Nuggets seem to have a questionably low calorie count - especially when we have a similar alternative from Burger King for comparison. Maybe it is because they are grilled rather than fried.

```{r, fig.width = 10, fig.height= 7}
min_cal_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(min_cal = min(calories))

min_cal_items <- fastfood %>% 
  inner_join(min_cal_numbers,
             by = c("calories" = "min_cal",
                    "restaurant"))

min_cal_items %>% 
  ggplot(aes(restaurant, calories))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))
```

# Items highest in cholesterol:

The recommended limit used to be 300mg per day or less. However, the current FDA recommendations no longer include this, as a [review of studies](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6024687/) has pointed out that cholesterol is not linked to cardiovascular disease.


```{r, fig.width = 10, fig.height= 7}

max_cholesterol_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_cholesterol = max(cholesterol))

max_cholesterol_items <- fastfood %>% 
  inner_join(max_cholesterol_numbers,
             by = c("cholesterol" = "max_cholesterol",
                    "restaurant"))

max_cholesterol_items %>% 
  ggplot(aes(restaurant, cholesterol))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_manual(values = wes_palette("Darjeeling1", type = "continuous", n = 8))

```

You know what *is* linked to cardiovascular disease, though?

# Items highest in saturated fat:

Guidelines recommend that people limit the calories from saturated and trans fats to no more than 10% of their caloric intake.

```{r, fig.width = 10, fig.height= 7}

max_sat_fat_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_sfat = max(sat_fat))

max_sat_fat_items <- fastfood %>% 
  inner_join(max_sat_fat_numbers,
             by = c("sat_fat" = "max_sfat",
                    "restaurant"))

max_sat_fat_items %>% 
  ggplot(aes(restaurant, sat_fat))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")+
  ylab("Saturated fat")

```

# Items highest in sodium:

The American Heart Association recommends [no more than 2,300 milligrams (mg) a day](https://www.heart.org/en/healthy-living/healthy-eating/eat-smart/sodium/how-much-sodium-should-i-eat-per-day) and moving toward an ideal limit of no more than 1,500 mg per day for most adults.

So, if you want to honor this, you'd better stay away from the following sodium bombs.

```{r, fig.width = 10, fig.height= 7}

max_sodium_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_sodium = max(sodium))

max_sodium_items <- fastfood %>% 
  inner_join(max_sodium_numbers,
             by = c("sodium" = "max_sodium",
                    "restaurant"))

max_sodium_items %>% 
  ggplot(aes(restaurant, sodium))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")

```

# Average sodium per item per brand: 

Arby's takes the lead, with McDonald's coming in second.

```{r, fig.width = 10, fig.height= 7}

fastfood %>% 
  group_by(restaurant) %>% 
  summarise(mean_sodium = mean(sodium)) %>% 
  ggplot(aes(reorder(restaurant, mean_sodium), mean_sodium))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = round(mean_sodium, 0)))+
  labs(x = "Brand",
       y = "Average sodium per item")+
  coord_flip()+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 8))
```

# Average calories per item per brand: 

McDonald's is the winner here. For us Lithuanians, Subway seems like a somewhat healthier alternative.

```{r, fig.width = 10, fig.height= 7}

fastfood %>% 
  group_by(restaurant) %>% 
  summarise(mean_calories = mean(calories)) %>% 
  ggplot(aes(reorder(restaurant, mean_calories), mean_calories))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = round(mean_calories, 0)))+
  labs(x = "Brand",
       y = "Average calories per item")+
  coord_flip()+
  scale_fill_tableau("Classic Cyclic")
```

# Let's introduce a couple of ratios: protein-to-calories and fiber-to-total-carbs

I have always wondered about the first one.

```{r, echo=TRUE}

fastfood <- fastfood %>% 
  mutate(
    protein_to_calorie = protein/calories,
    fiber_to_carb = fiber/total_carb
  )

```

Foods that offer the most protein per calorie:

```{r, fig.width = 10, fig.height= 7}

'%!in%' <- function(x,y)!('%in%'(x,y))

max_ptc_numbers <- fastfood %>%
  group_by(restaurant) %>% 
  summarise(max_ptc = max(protein_to_calorie))

max_ptc_items <- fastfood %>% 
  inner_join(max_ptc_numbers,
             by = c("protein_to_calorie" = "max_ptc",
                    "restaurant"))

max_ptc_items %>% 
  filter(restaurant %!in% c("Burger King")) %>% 
  ggplot(aes(restaurant, protein_to_calorie))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")+
  labs(x = "Brand",
       y = "Protein to calorie ratio")

```

Foods with the highest fiber-to-carbohydrate ratio: I had to leave only some of the brands here due to issues like missing data.

```{r}

max_ftc_numbers <- fastfood %>%
  group_by(restaurant) %>% 
  summarise(max_ftc = max(fiber_to_carb))

max_ftc_items <- fastfood %>% 
  inner_join(max_ftc_numbers,
             by = c("fiber_to_carb" = "max_ftc",
                    "restaurant"))

max_ftc_items %>% 
  filter(restaurant %in% c("Sonic", "Subway", "Mcdonalds")) %>% 
  ggplot(aes(restaurant, fiber_to_carb))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
   scale_fill_manual(values = wes_palette("GrandBudapest1",  n = 3))+
  labs(x = "Brand",
       y = "Fiber to carbohydrate ratio")

```

# McDonald's vs Burger King: a battle

The top 10 items highest in calories for both brands: McDonald's takes 6 of these.

```{r, fig.width = 10, fig.height= 7}

litbrands <- fastfood %>% 
  filter(restaurant %in% c("Mcdonalds", "Burger King"))

litbrands %>% 
  group_by(restaurant) %>% 
  arrange(desc(calories)) %>% 
  head(10) %>% 
  ggplot(aes(reorder(item,calories), calories))+
  geom_col(aes(fill = item),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label =  restaurant))+
  coord_flip()+
  xlab("Item")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 10))

```

When we look at the top 10 items for both brands ranked by the protein-to-calorie ratio, McDonald's takes an obvious lead, claiming eight of the places.

```{r, fig.width = 10, fig.height= 7}

litbrands %>% 
  group_by(restaurant) %>% 
  arrange(desc(protein_to_calorie)) %>% 
  head(10) %>% 
  ggplot(aes(reorder(item,protein_to_calorie), protein_to_calorie))+
  geom_col(aes(fill = item),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label =  restaurant))+
  coord_flip()+
  labs(x = "Item", 
       y = "Protein-to-calorie ratio")+
   scale_fill_manual(values = wes_palette("Darjeeling1", type = "continuous", n = 10))

```

What if we looked at protein only, disregarding the calories?

Even if you're not aiming to maintain a caloric deficit, McDonald's is still superior when it comes to offering high-protein items.

```{r, fig.width = 10, fig.height= 7}
litbrands %>% 
  group_by(restaurant) %>% 
  arrange(desc(protein)) %>% 
  head(10) %>% 
  ggplot(aes(reorder(item,protein), protein))+
  geom_col(aes(fill = item),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label =  restaurant))+
  coord_flip()+
  labs(x = "Item", 
       y = "Protein")+
   scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 10))
```

# Looking at the McDonald's highest protein-to-calorie item: 

The Premium Bacon Ranch Salad with Grilled Chicken contains an impressive ratio of protein.

```{r}

m <- fastfood %>% 
  filter(X1 == 53)%>%
  select(-c(X1, restaurant, salad)) %>% 
  pivot_longer(cols = -item,
               names_to = "nutrient",
               values_to = "quantity"
               )

m %>% 
  filter(nutrient %in% c("protein", "total_fat", "total_carb")) %>% 
  mutate(nutrient = factor(nutrient),
         nutrient = fct_recode(nutrient,
           `Protein` = "protein",
           `Total Fat` = "total_fat",
           `Total Carbohydrates` = "total_carb"
         )) %>% 
  ggplot(aes(x = "", y = quantity, fill = nutrient))+
  geom_bar(width = 1, stat = "identity",
           color = "black")+
  coord_polar("y", start=0)+
  theme_void()+
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 3))+
  ylab("Composition of the Premium Bacon Ranch Salad")+
  theme(panel.background = element_rect(fill = "#fff2cf", color = "#fff2cf"))

```