---
title: "FastFood"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}

library(readr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(dplyr)
library(wesanderson)
library(ggthemes)

fastfood <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-09-04/fastfood_calories.csv')
theme_set(theme_solarized())


```

# Brands by item count:

```{r, fig.width = 10, fig.height= 7}
fastfood %>% 
  group_by(restaurant) %>% 
  summarise(total_items = sum(n())) %>% 
  ggplot()+
  geom_col(aes(reorder(restaurant, total_items), total_items, fill = restaurant),
           show.legend = FALSE, color = "black")+
  labs(x ="Restaurant",
       y ="Total item count")+
  scale_fill_manual(values = wes_palette("Zissou1", type = "continuous", n = 8))
```

# Highest calorie item per brand:

```{r, fig.width = 10, fig.height= 7}
max_cal_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_cal = max(calories))

max_cal_items <- fastfood %>% 
  inner_join(max_cal_numbers,
             by = c("calories" = "max_cal",
                    "restaurant"))

max_cal_items %>% 
  ggplot(aes(restaurant, calories))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")

```

# Lowest calorie item per brand:

```{r, fig.width = 10, fig.height= 7}
min_cal_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(min_cal = min(calories))

min_cal_items <- fastfood %>% 
  inner_join(min_cal_numbers,
             by = c("calories" = "min_cal",
                    "restaurant"))

min_cal_items %>% 
  ggplot(aes(restaurant, calories))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))
```

# Items highest in cholesterol:

```{r, fig.width = 10, fig.height= 7}

max_cholesterol_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_cholesterol = max(cholesterol))

max_cholesterol_items <- fastfood %>% 
  inner_join(max_cholesterol_numbers,
             by = c("cholesterol" = "max_cholesterol",
                    "restaurant"))

max_cholesterol_items %>% 
  ggplot(aes(restaurant, cholesterol))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_manual(values = wes_palette("Darjeeling1", type = "continuous", n = 8))

```

# Items highest in sodium:

```{r, fig.width = 10, fig.height= 7}

max_sodium_numbers <- fastfood %>% 
  group_by(restaurant) %>% 
  summarise(max_sodium = max(sodium))

max_sodium_items <- fastfood %>% 
  inner_join(max_sodium_numbers,
             by = c("sodium" = "max_sodium",
                    "restaurant"))

max_sodium_items %>% 
  ggplot(aes(restaurant, sodium))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")

```

# Average calories per item per brand:

```{r}

fastfood %>% 
  group_by(restaurant) %>% 
  summarise(mean_calories = mean(calories)) %>% 
  ggplot()+
  geom_col(aes(reorder(restaurant, mean_calories), mean_calories, fill = restaurant),
           show.legend = FALSE, color = "black")+
  labs(x = "Brand",
       y = "Average calories per item")+
  coord_flip()
```

# Let's introduce a couple of ratios: protein-to-calories and fiber-to-total-carbs

```{r, echo=TRUE}

fastfood <- fastfood %>% 
  mutate(
    protein_to_calorie = protein/calories,
    fiber_to_carb = fiber/total_carb
  )

```

Foods that offer the most protein per calorie:

```{r}

'%!in%' <- function(x,y)!('%in%'(x,y))

max_ptc_numbers <- fastfood %>%
  group_by(restaurant) %>% 
  summarise(max_ptc = max(protein_to_calorie))

max_ptc_items <- fastfood %>% 
  inner_join(max_ptc_numbers,
             by = c("protein_to_calorie" = "max_ptc",
                    "restaurant"))

max_ptc_items %>% 
  filter(restaurant %!in% c("Burger King")) %>% 
  ggplot(aes(restaurant, protein_to_calorie))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")+
  labs(x = "Brand",
       y = "Protein to calorie ratio")

```

Foods with the highest fiber-to-carbohydrate ratio:

```{r}

max_ftc_numbers <- fastfood %>%
  group_by(restaurant) %>% 
  summarise(max_ftc = max(fiber_to_carb))

max_ftc_items <- fastfood %>% 
  inner_join(max_ftc_numbers,
             by = c("fiber_to_carb" = "max_ftc",
                    "restaurant"))

max_ftc_items %>% 
  filter(restaurant %in% c("Sonic", "Subway", "Mcdonalds")) %>% 
  ggplot(aes(restaurant, fiber_to_carb))+
  geom_col(aes(fill = restaurant),
           show.legend = FALSE, color = "black")+
  geom_label(aes(label = item),
             label.size = 0.05,
             label.padding = unit(0.2, "lines"))+
  scale_fill_tableau("Classic Cyclic")+
  labs(x = "Brand",
       y = "Fiber to carbohydrate ratio")

```

# A closer look into the brands available (or soon-to-be-available) in Lithuania:

That would be McDonald's, Subway and Burger King.

```{r}

litbrands <- fastfood %>% 
  filter(restaurant %in% c("Mcdonalds", "Subway", "Burger King"))

```

# Looking at the McDonald's highest protein-to-calorie item:

```{r}

m <- fastfood %>% 
  filter(X1 == 53)%>%
  select(-c(X1, restaurant, salad)) %>% 
  pivot_longer(cols = -item,
               names_to = "nutrient",
               values_to = "quantity"
               )

m %>% 
  filter(nutrient %in% c("protein", "total_fat", "total_carb")) %>% 
  mutate(nutrient = factor(nutrient),
         nutrient = fct_recode(nutrient,
           `Protein` = "protein",
           `Total Fat` = "total_fat",
           `Total Carbohydrates` = "total_carb"
         )) %>% 
  ggplot(aes(x = "", y = quantity, fill = nutrient))+
  geom_bar(width = 1, stat = "identity",
           color = "black")+
  coord_polar("y", start=0)+
  theme_void()+
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 3))+
  ylab("Composition of the Premium Bacon Ranch Salad")+
  theme(panel.background = element_rect(fill = "#fff2cf", color = "#fff2cf"))

```